---
title: "Sampling strategies for age composition in commercial sampling and its effects on age distribution estimates"
author: "Laurent Dubroca in WKBIOPTIM3"
date: "`r format(Sys.time(),'%d/%m/%Y')`"
output: 
  pdf_document
bibliography: '/home/moi/datahome/work/biblio/enfin/biblioloran.bib'
---

# Introduction

Some words about the framework, the needs and etc.
This document follows the framework proposed by @aanes2015.


```{r global_options, include=T,cache=F,echo=F,warning=F,message=F,progress=F,verbose=F}
#knitr option
operationnel<-TRUE
knitr::opts_chunk$set(echo=FALSE, 
		      warning=!operationnel, 
		      message=!operationnel,
		      fig.height=4,
		      progress=!operationnel,
		      verbose=!operationnel,
		      include=TRUE,dev='png',autodep=FALSE)
#package
library(dplyr)
library(ggplot2)
library(sampling)
library(FSA)
library(FSAsim)
library(purrr)
##library(openxlsx);library(mailR)
library(pander);library(captioner)
#local fct
#source("credo_fct.R")
#initialise les fonctions pour légender tables et figures
tabcap <- captioner(prefix = "Table")
figcap <- captioner(prefix = "Figure")

#general parameters
nbpopfish<-100000 #number of fish in the virtual population
natmortpopfish<-0.3 # the natural mortality of the virtual population
vbLinf <- 500  #von berta: Linf value of the virtual fish
vbK <- 0.3#von berta: K value of the virtual fish
vbt0 <- 0#von berta: t0 of the virtual fish
vbsigma <- 20#von berta sd around age for length distribution 

#fix randomness with the number of the Beast
set.seed(666)
```

# Materials and methods


## Simulation

Simulation of a virtual fish population.
Principles and paramaters.


### A virtual fish population 

Comments of `r figcap("popsimplt1",display="cite")` and `r figcap("popsimplt2",display="cite")`.


```{r popsim1, include=T,cache=T,echo=F,eval=T}
#a population of fish
popage<-simAges(N0=nbpopfish,A=natmortpopfish,incl.zero=T)
#their lengths
poplen <- simLenFromAge(popage,Linf=vbLinf,K=vbK,t0=vbK,sigma=vbsigma)
#the population dataframe (filtering out negative length if any) 
popfish <- data.frame(age=popage,len=poplen)%>%filter(len>0)
#trunc the fish size
popfish$len <- lencat(popfish$len,w=10)
#some text summarizing the simulation parameters
txtparam1<-paste0("n=",nbpopfish,",A=",natmortpopfish)
txtparam2<-paste0("Linf=",vbLinf,",K=",vbK,",t0=",vbt0,",sd=",vbsigma)
```

```{r popsimplt1, include=T,cache=T,echo=F,eval=T,height=5}
p1<-ggplot(popfish,aes(x=age,y=len))+geom_point(alpha=.4)+
	theme_bw()+xlab("Age (year)")+ylab("Length (mm)")+
	ggtitle(paste0("Length at age of the simulated population\n",txtparam1,"\n",txtparam2))
print(p1)
```

```{r popsimplt1cap,cache=T,eval=T,echo=FALSE,results="asis"}
cat(figcap(name="popsimplt1",caption="Length at age in the virtual population"))
```


```{r popsimplt2, include=T,cache=T,echo=F,eval=T,height=5}
p2<-ggplot(gather(popfish),aes(value))+geom_histogram()+#binwidth=c(1,10))+
	facet_wrap(~key,scale="free")+
	theme_bw()+ylab("Count")+xlab("Value")+
	ggtitle(paste0("Age and length distribution of the simulated population\n",txtparam1,"\n",txtparam2))
print(p2)
```
```{r popsimplt2cap,cache=T,eval=T,echo=FALSE,results="asis"}
cat(figcap(name="popsimplt2",caption="Length distribution of the virtual population"))
```


### A virtual fleet targeting a virtual fish population 

Principles.
A gear is define by its selectiviy curve. This curve is supplied using a beta
distribution.
An example of a selectivity curve

```{r fleetexample, include=T,cache=T,echo=F,eval=T}
selex1<-data.frame(x=seq(0,1,0.01))%>%mutate(sel=dbeta(x,shape1=10,shape2=20),gear="1")
selex2<-data.frame(x=seq(0,1,0.01))%>%mutate(sel=dbeta(x,shape1=4,shape2=3),gear="2")
selex3<-data.frame(x=seq(0,1,0.01))%>%mutate(sel=dbeta(x,shape1=8,shape2=1),gear="3")
selex<-rbind(selex1,selex2,selex3)
p1<-ggplot(selex,aes(x=x,y=sel,color=gear))+
	geom_path()+
	theme_bw()+ylab("Selectity")+xlab("Normalized size")+
	ggtitle(paste0("Selectivity curve from three gamma distributions"))#\n",txtparam1,"\n",txtparam2))
print(p1)
```

Population by gears + trips: 

```{r fleetandtrip, include=T,cache=T,echo=F,eval=T}
#population accessible to each trip
lengear1<-simLenSelectP(popfish$len, alpha = 10, beta = 20, max.height = 1, show = FALSE)
lengear2<-simLenSelectP(popfish$len, alpha = 4, beta = 3, max.height = 1, show = FALSE)
lengear3<-simLenSelectP(popfish$len, alpha = 8, beta = 1, max.height = 1, show = FALSE)
popfish1<-popfish%>%filter(lengear1)%>%mutate(gear="1")
popfish2<-popfish%>%filter(lengear2)%>%mutate(gear="2")
popfish3<-popfish%>%filter(lengear3)%>%mutate(gear="3")
popfishgear<-rbind(popfish1,popfish2,popfish3)
p1<-ggplot(gather(popfishgear,key="key",value="value",age:len),
	   aes(value,fill=gear,group=gear))+
	geom_histogram(position="dodge")+
	facet_wrap(~key,scale="free")+
	theme_bw()+ylab("Count")+xlab("Value")+
	ggtitle(paste0("Age and length distributions available by gear"))
print(p1)

#generate some trips
#a fct to generate 1 trip
do1trip<-function(df){sample_n(df,300)%>%mutate(id=sub("\\.","",format(Sys.time(),"%s%OS3")))}
#generate 100 trip for popfish1
trip1<-do.call("rbind",rerun(100,do1trip(popfish1))) %>%mutate(gear="1")
trip2<-do.call("rbind",rerun(100,do1trip(popfish2))) %>%mutate(gear="2") 
trip3<-do.call("rbind",rerun(100,do1trip(popfish3))) %>%mutate(gear="3")
tripgear<-rbind(trip1,trip2,trip3)

#length distribution of the trips by gear
p1<-ggplot(gather(tripgear,key="key",value="value",age:len),
	   aes(value,fill=gear,group=gear))+
	#geom_histogram(alpha=1,position="dodge",binwidth=10)+
	geom_histogram(alpha=1,position="dodge")+
	facet_wrap(~key,scale="free")+
	theme_bw()+ylab("Count")+xlab("Value")+
	ggtitle(paste0("Age and length distributions of the trips by gear"))
print(p1)
```

## ALK sampling

Samples of ALK.
- samples the trips,
- independent samples,
- ??

```{r alk, include=T,cache=T,echo=F,eval=F}
ranalk<-function(nb=100){
# random sample of the pop
	sampalk<-sample_n(popfish,100)
	#complete the alkprop with zero for the non sample length
	lenall<-data.frame(len=sort(unique(popfishgear$len)))
	sampalkall<-left_join(lenall,sampalk,by="len")#%>%
	sampalkall[is.na(sampalkall)]<-0
	#compute the alk
	alk <- xtabs(~len+age,data=sampalkall)#%>%transmute(age=trueage,lenCat))
	alkprop <- prop.table(alk,margin=1)
	rez1<-alkIndivAge(alkprop,age~len,data=popfishgear)#%>%mutate(trueage=age))
	rez1%>%mutate(id="ranalk",nb,rep=sub("\\.","",format(Sys.time(),"%s%OS3")))
}


#a graph ?
rez1<-ranalk(100)
do1trip<-function(df){sample_n(df,300)%>%mutate(rep=sub("\\.","",format(Sys.time(),"%s%OS3")))}
#generate 100 trip for popfish1
ranalk100<-do.call("rbind",rerun(100,ranalk(100)))



pipo<-rbind(rez1%>%select(age)%>%mutate(type="estim"),
	    popfishgear%>%select(age)%>%mutate(type="pop"))
p1<-ggplot(pipo,aes(age,fill=type,group=type))+
	geom_histogram(alpha=1,position="dodge",binwidth=1)+
	theme_bw()+ylab("")+xlab("Length (mm)")+
	ggtitle(paste0("Length distribution of the trips by gear"))
print(p1)

earthmovdist::emdL1(rez1$age,popfishgear$age)


```



```{r data_and_param000, include=T,cache=T,echo=F,eval=F}





#apply age length key to length data
#from FSA
rez1<-alkIndivAge(alkprop,age~lenCat,data=df1%>%mutate(age=0))
rez1hist<-rez1%>%group_by(age)%>%summarise(n=n()*2,type="sampalk")
rez2hist<-rez1%>%group_by(age=trueage)%>%summarise(n=n()*2,type="sampreal")
rez3hist<-df%>%group_by(age)%>%summarise(n=n(),type="pop")
rez1hist<-rbind(rez1hist,rez2hist,rez3hist)
ggplot(rez1hist,aes(x=age,y=n,color=type))+geom_point()+geom_path()

#a more complete example
#a population of fish
tabage<-simAges(N0=100000,A=.35,incl.zero=T)
#their lengths
tablen <- simLenFromAge(tabage,Linf=228,K=0.206,t0=0,sigma=8)
## Some summaries
df <- data.frame(age=tabage,len=tablen)%>%filter(len>0)
Summarize(len~age,data=df,digits=1)
plot(len~age,data=df,pch=16,col=rgb(0,0,0,0.1),xlab="Age",ylab="Length")
hist(df$len,n=50)





# independant sample
nbbyage<-rep(5,length(unique(df1$lenCat)))
df2<-strata(df1,stratanames="lenCat",size=nbbyage,method="srswr")
df2$trueage<-df1$trueage[df2$ID_unit]
alk <- xtabs(~lenCat+age,data=df2%>%transmute(age=trueage,lenCat))
alkprop <- prop.table(alk,margin=1)
round(alkprop,2)

```



```{r data_and_param, include=T,cache=T,echo=F,eval=F}





#apply age length key to length data
#from FSA
rez1<-alkIndivAge(alkprop,age~lenCat,data=df1%>%mutate(age=0))
rez1hist<-rez1%>%group_by(age)%>%summarise(n=n()*2,type="sampalk")
rez2hist<-rez1%>%group_by(age=trueage)%>%summarise(n=n()*2,type="sampreal")
rez3hist<-df%>%group_by(age)%>%summarise(n=n(),type="pop")
rez1hist<-rbind(rez1hist,rez2hist,rez3hist)
ggplot(rez1hist,aes(x=age,y=n,color=type))+geom_point()+geom_path()

#a more complete example
#a population of fish
tabage<-simAges(N0=100000,A=.35,incl.zero=T)
#their lengths
tablen <- simLenFromAge(tabage,Linf=228,K=0.206,t0=0,sigma=8)
## Some summaries
df <- data.frame(age=tabage,len=tablen)%>%filter(len>0)
Summarize(len~age,data=df,digits=1)
plot(len~age,data=df,pch=16,col=rgb(0,0,0,0.1),xlab="Age",ylab="Length")
hist(df$len,n=50)
#1 gear no selectivity
## take some length sample randomly
df1<-df%>%sample_frac(0.01)
df1$lenCat <- lencat(df1$len,w=10)
df1<-df1%>%mutate(trueage=age)%>%select(-age)
## sample age based on df1
fct1<-function(df1,nb=5){
	nbbyage<-rep(nb,length(unique(df1$lenCat)))
	df2<-strata(df1,stratanames="lenCat",size=nbbyage,method="srswr")
	df2$trueage<-df1$trueage[df2$ID_unit]
	alk <- xtabs(~lenCat+age,data=df2%>%transmute(age=trueage,lenCat))
	alkprop <- prop.table(alk,margin=1)
	rez1<-alkIndivAge(alkprop,age~lenCat,data=df1%>%mutate(age=0))
	rez1%>%mutate(n=nb)
}
#generate differente sample rate
rezall<-data.frame()
for(i in 1:30){
	rez0<-fct1(df1,i)
	rezall<-rbind(rez0,rezall)
}
#apply age length key to length data
#from FSA
prop<-1/0.01
rez1hist<-rezall%>%group_by(age,nbsamp=n)%>%summarise(n=n()*prop,type="sampalk")%>%ungroup()
rez2hist<-df1%>%group_by(age=trueage)%>%summarise(n=n()*prop,type="sampreal")%>%ungroup()
rez3hist<-df%>%group_by(age)%>%summarise(n=n(),type="pop")%>%ungroup()
#rez1hist<-rbind(rez1hist,rez2hist,rez3hist)
#ggplot(rez1hist,aes(x=age,y=n,color=nbsamp))+geom_boxplot()+
ggplot(data=rez1hist,aes(x=age,y=n,group=age))+geom_boxplot()+
geom_point(data=rez1hist,aes(x=age,y=n,color=nbsamp,group=NULL),alpha=.2)+#geom_point(alpha=.2)+
	geom_line(data=rez2hist,aes(x=age,y=n,group=NULL),col="red")+
	geom_path(data=rez3hist,aes(x=age,y=n,group=NULL),col="green")

#######################################################
#a more complete example, again, with PSU and so on
#a population of fish
tabage<-simAges(N0=100000,A=.35,incl.zero=T)
#their lengths
tablen <- simLenFromAge(tabage,Linf=228,K=0.206,t0=0,sigma=8)
## Some summaries
df <- data.frame(age=tabage,len=tablen)%>%filter(len>0)
Summarize(len~age,data=df,digits=1)
plot(len~age,data=df,pch=16,col=rgb(0,0,0,0.1),xlab="Age",ylab="Length")
hist(df$len,n=50)
dfhist<-df%>%group_by(len)%>%summarise(n=n(),type="pop")

#1 gear with a given selectivity
aa<-simLenSelectP(df$len, alpha = 3, beta = 10, max.height = 1, show = FALSE)
df1<-df%>%filter(aa)%>%sample_frac(0.01)
dfhist1<-df1%>%group_by(len)%>%summarise(n=n(),type="gear1")
ggplot(rbind(dfhist,dfhist1),aes(x=len,y=n,col="type"))+geom_path()+facet_wrap(~type,scale="free")

#generate xx trips
library(purrr)
dontrip<-function(df,n=5,prop0=0.01,alpha0=5,beta0=10){
	dotrip<-function(df,prop=prop0,alpha=alpha0,beta=beta0){
		aa<-simLenSelectP(df$len, alpha, beta, max.height = 1, show = FALSE)
		df%>%filter(aa)%>%sample_frac(0.01)%>%
			mutate(trip=paste(sample(letters,5),collapse=""))
	}
	aa<-rerun(n,dotrip(df,prop=prop0,alpha=alpha0,beta=beta0))
	aa<-do.call("rbind",aa)
	aa$lenCat <- lencat(aa$len,w=10)
	return(aa)
}

trip1<-dontrip(df,n=10,alpha0=2,beta0=7)
trip1hist<-trip1%>%group_by(len,trip)%>%summarise(n=n(),type="gear1")
prop<-max(trip1hist$n)/max(dfhist$n)
ggplot(trip1hist,aes(x=len,y=n,col=trip))+geom_path()+
	geom_path(data=dfhist,aes(x=len,y=n*prop,col="type"),col="black")

trip2<-dontrip(df,n=10,alpha0=6,beta0=7)
trip1hist<-trip2%>%group_by(len,trip)%>%summarise(n=n(),type="gear1")
prop<-max(trip1hist$n)/max(dfhist$n)
ggplot(trip1hist,aes(x=len,y=n,col=trip))+geom_path()+
	geom_path(data=dfhist,aes(x=len,y=n*prop,col="type"),col="black")

## sample age based on df1
fct1<-function(df1=rbind(trip1,trip2),nb=5){
df1=rbind(trip1,trip2);nb<-5
	nbbyage<-rep(nb,length(unique(df1$lenCat)))
	df2<-strata(df1,stratanames="lenCat",size=nbbyage,method="srswr")
	df2$trueage<-df1$age[df2$ID_unit]
	alk <- xtabs(~lenCat+age,data=df2%>%transmute(age=trueage,lenCat))
	alkprop <- prop.table(alk,margin=1)
	rez1<-alkIndivAge(alkprop,age~lenCat,data=df1%>%mutate(age=0))
	rez1%>%mutate(n=nb)
}

#generate differente sample rate
rezall<-data.frame()
for(i in 1:30){
	rez0<-fct1(df1,i)
	rezall<-rbind(rez0,rezall)
}

#apply age length key to length data
#from FSA
prop<-1/0.01
rez1hist<-rezall%>%group_by(age,nbsamp=n)%>%summarise(n=n()*prop,type="sampalk")%>%ungroup()
rez2hist<-df1%>%group_by(age=trueage)%>%summarise(n=n()*prop,type="sampreal")%>%ungroup()
rez3hist<-df%>%group_by(age)%>%summarise(n=n(),type="pop")%>%ungroup()
#rez1hist<-rbind(rez1hist,rez2hist,rez3hist)
#ggplot(rez1hist,aes(x=age,y=n,color=nbsamp))+geom_boxplot()+
ggplot(data=rez1hist,aes(x=age,y=n,group=age))+geom_boxplot()+
geom_point(data=rez1hist,aes(x=age,y=n,color=nbsamp,group=NULL),alpha=.2)+#geom_point(alpha=.2)+
	geom_line(data=rez2hist,aes(x=age,y=n,group=NULL),col="red")+
	geom_path(data=rez3hist,aes(x=age,y=n,group=NULL),col="green")






```

# References

