---
title: 'WKBIOPTIM2: optimization at the population level'
author: "John Doe"
date: '`r date()`'
output:
  html_document: default
---

This report define the simulation scenario.


```{r compil,eval=F,include=F,cache=F,echo=F,warning=F,message=F,progress=F,verbose=F}
#just a reminder to how compile a docx doc
if(F){
rmarkdown::render("",c("word_document"))
}
```


#Tools

R, Rstudio and packages.

```{r global_options}
#working directory have to be the directory where is the Rmarkdown file.
setwd("~/ifremer/wkbioptim2/scripts/pop_level_sim_ope")
#R general option:
options(stringsAsFactors = FALSE)
#chunk option
knitr::opts_chunk$set(cache=FALSE,echo=TRUE, warning=FALSE, 
	message=FALSE, fig.height=6,progress=FALSE,verbose=FALSE,
		include=TRUE,dev='png',autodep=FALSE)
#package
library(COSTcore)
library(COSTdbe)
library(COSTeda)
library(dplyr) 
library(ggplot2)
library(maps)
library(mapdata)
library(pander)
#DATRAS is not available on CRAN so use devtools:
#devtools::install_github("DTUAqua/DATRAS/DATRAS")
library(DATRAS)
#library(earthmovdist) deprecated
library(emdist)
library(sampling) # for stratified random sampling

#source local functions file
source("fct.R")

#load the data
load("./data/dataready.rdata")

```
#Data

The data saved in the previous analyses are loaded.

```{r data1,eval=F}
load("./data/dataready.rdata")
```

#Sampling object randomly

The main process is sample removal in the original data.
In `CSc` object, the variable `trpCode` is the trip identifier, while staNum is
the haul id (unique *only* by `trpcode`).

##Trip selection

###Random trip selection

n trips are selected randomly with replacement.
In the following example `n=20`.

```{r randomtrip}
	idtrip<-unique(CSc@hh$trpCode)
	idtriprand<-sample(idtrip,20,replace=T)	
```

###Stratified random trip selection

n trips are selected randomly with replacement stratified by quarter.
In the following example `n=20`.

```{r stratrandomtrip1}
	idtrip<-CSc@hh%>%select(trpCode,strata=time)%>%distinct()
	nbbystrat<-rep(20,length(unique(idtrip$strata)))
	idtripstrata<-strata(idtrip,stratanames="strata",size=nbbystrat,method="srswr")
	idtripstrata$trpCode<-idtrip$trpCode[idtripstrata$ID_unit]
```

n trips are selected randomly with replacement stratified by quarter and metier.
In the following example `n=20`.

```{r stratrandomtrip2}
	idtrip<-CSc@hh%>%
		group_by(trpCode)%>%
		summarize(strata1=unique(time),strata2=paste(sort(unique(technical)),collapse=","))%>%
		ungroup()
	nbbystrat<-rep(20,length(paste(idtrip$strata1,idtrip$strata2)))
	idtripstrata<-strata(idtrip,stratanames=c("strata1","strata2"),size=nbbystrat,method="srswr")
	idtripstrata$trpCode<-idtrip$trpCode[idtripstrata$ID_unit]
```

##Haul selection

###Random haul selection

n hauls are selected randomly with replacement.
In the following example `n=20`.

```{r randomhaul}
	idhaul<-unique(paste(CSc@hh$trpCode,CSc@hh$staNum))
	idhaulrand<-sample(idhaul,20,replace=T)	
```

Stratified haul selection can be done following the trips example.


#Simulation

Use the functions `simutrip` and `simutriplan`

##Trips

```{r simurandomtrip}
	rez<-simutrip(CSc,myStr,c(5,10,20),3)
```

```{r simurandomtripplt}
	ggplot(na.omit(rez[[1]]),aes(x=length,y=value,color=n))+
		 scale_color_distiller(palette='Spectral')+
		geom_point(alpha=.2)+
		facet_grid(spp~technical,scale="free_y")+
		theme_bw()+
 		theme(axis.text.x = element_text(size=8, angle=90),
		axis.text.y = element_text(size=10, angle=0),
		strip.text.x=element_text(size=8,angle=90),
		strip.text.y=element_text(size=8,angle=0),
		legend.position="bottom")
	ggplot(na.omit(rez[[2]]),aes(x=length,y=value,color=n))+
		 scale_color_distiller(palette='Spectral')+
		geom_point(alpha=.2)+
		facet_grid(spp~technical,scale="free_y")+
		theme_bw()+
 		theme(axis.text.x = element_text(size=8, angle=90),
		axis.text.y = element_text(size=10, angle=0),
		strip.text.x=element_text(size=8,angle=90),
		strip.text.y=element_text(size=8,angle=0),
		legend.position="bottom")
	ggplot(na.omit(rez[[3]]),aes(x=n,y=wcv))+
		geom_poin()+
		facet_grid(~spp)

```



