---
title: "SDTool (STREAM project: (STrengthening REgional cooperation in the Area of fisheries biological data collection in the Mediterranean and Black Sea)): Investigate Data"
author: "M.T. Facchini, I. Bitetto (COISPA, Italy)"
date: '`r date()`'
output: word_document
---


```{r compil, eval=F,include=F,cache=F,echo=F,warning=F,message=F,progress=F,verbose=F}

# SD Tool v.3 and WKBIOPTIM SimTool v.1  were taken into account for the developement of the present tool.
# SD tool has been developed in MARE/2014/19 project and updated in MARE/2016/22 project (STREAM)

# The tool is documented in Deliverable D3.3 "Upgrade the methodological framework and tools for sampling optimization, implement and report case studies" (January 2019)

# Authors of the first version of SD Tool: F. Gontrand, T. Rouyer, N. Billet, 2016
# IFREMER, UMR MARBEC, Avenue Jean Monnet, BP 171, 34203 S` ete, France 

# Authors of the first version of WKBIOPTIM scripts: Nuno Prista (SLU, Sweden) from a subgroup work carried out during a local Workshop on Sampling Design and Optimization (Lysekil, 31 October to 4 November 2016, unpublished)
# SLU Aqua, Institute of Marine Research, Department of Acquatic Resources - Swedish University of Agricultural Sciences

# Authors of this new tool: M.T. Facchini, I. Bitetto, 2017
# COISPA Tecnologia & Ricerca, Via dei Trulli 18/20 - (Bari), Italy 

# In case of use of the tool, the Authors should be cited.
# If you have any comments or suggestions please contact the following e-mail addresses: facchini@coispa.it, bitetto@coispa.it
# SD Tool is believed to be reliable. However, we disclaim any implied warranty or representation 
# about its accuracy, completeness or appropriateness for any particular purpose.

```


#Tools

R, Rstudio and packages.


```{r results='hide', message=FALSE, warning=FALSE}


#chunk option
knitr::opts_chunk$set(cache=FALSE,echo=TRUE, warning=FALSE, 
	message=FALSE, fig.height=6,progress=FALSE,verbose=FALSE,
		include=TRUE,dev='png',autodep=FALSE)

#load packages
library(fishPifct)
library(pander)
library(COSTcore)
library(COSTdbe)
library(COSTeda)
library(data.table) 

#pander options
panderOptions('table.split.table', 60)
panderOptions('table.style', 'grid')
panderOptions('table.split.cells', 10)
panderOptions('table.alignment.default', 'left')
panderOptions('table.alignment.rownames', 'right')
panderOptions('decimal.mark', ',')
panderOptions('graph.fontsize', '10')


```



# fishPifct code for the SDEF format check

fishPifct is an R package combining some functions used in the fishPi project (fishPi). Some of these functions are coming from the SDEFQuality package from Norbert Billet (SDEFQuality).
For further details visit the GitHub web page at the following link: https://github.com/ldbk/fishPifct 
It is suggested to use it before running the analysis.


##Settings 

```{r settings}

# set the working directory
# set the directory for saving results
myWD <- "C:\\Users\\Bitetto Isabella\\OneDrive - Coispa Tecnologia & Ricerca S.C.A.R.L\\MARE22\\STREAM\\FINAL REVISION OF DELIVERABLES\\To upload on GITHUB\\WP3\\Task 3.3\\SD tool 2.04"
setwd(myWD)

results_path <- "C:\\Users\\Bitetto Isabella\\OneDrive - Coispa Tecnologia & Ricerca S.C.A.R.L\\MARE22\\STREAM\\FINAL REVISION OF DELIVERABLES\\To upload on GITHUB\\WP3\\Task 3.3\\SD tool 2.04" 

sapply(list.files(pattern="[.]R$", path="SDTfuns/", full.names=TRUE), source);
  
cl_all <- readRDS(paste(results_path, "/input files/costCL.rds", sep=""))
cs_all <- readRDS(paste(results_path, "/input files/costCS.rds", sep=""))

metier_shortcode <<- read.csv(paste(results_path, "/input files/metier_shortcode.csv", sep=""), sep=";")
 
table_strat_res <<- read.csv(paste(results_path, "/input files/05_1_stratification_results.csv", sep=""), sep=";")

years <- unique(cs_all@tr$year)
years <- years[order(years)]

 target_SPECIES <<- list("DPS" = "Parapenaeus longirostris")

gsas <- unique(cs_all@hh$area)

head(cs_all@tr)
head(cs_all@hh)
head(cs_all@sl)
head(cs_all@hl)
head(cs_all@ca)

# strDates <- cs_all@hh$date
# dates <- as.Date(strDates, "%d/%m/%Y")
# cs_all@hh$date <- dates
  
cs_all@hh$foVal <- "V"

cs_all@sl$landCat <- "HUC"
cs_all@hl$landCat <- "HUC"

cs_all@sl$subSampCat  <- ""
cs_all@hl$subSampCat  <- ""

cs_all@sl$catchCat <- toupper(cs_all@sl$catchCat) 
cs_all@hl$catchCat <- toupper(cs_all@hl$catchCat) 
cs_all@ca$catchCat <- toupper(cs_all@ca$catchCat)


dir.create(results_path)

for (nr in 1:nrow(cl_all@cl)) {
  if (length(as.character(metier_shortcode$lev4[as.character(metier_shortcode$Fishing_activity_category_level6) == cl_all@cl$foCatEu6[nr]]) ) > 0) {
       cl_all@cl$foCatEu5[nr] =  as.character(metier_shortcode$lev4[as.character(metier_shortcode$Fishing_activity_category_level6) == cl_all@cl$foCatEu6[nr]])
  cl_all@cl$foCatEu6[nr] = as.character(metier_shortcode$short_code[as.character(metier_shortcode$Fishing_activity_category_level6) == cl_all@cl$foCatEu6[nr]]) 
  
} else {
  print(paste("NOT FOUND: ", cl_all@cl$foCatEu6[nr]))
}
} 
  
   for (nr in 1:nrow(cs_all@hh)) {
   if ( length(as.character(metier_shortcode$lev4[as.character(metier_shortcode$Fishing_activity_category_level6) == cs_all@hh$foCatEu6[nr]]) ) >0) {
cs_all@hh$foCatEu5[nr] =  as.character(metier_shortcode$lev4[as.character(metier_shortcode$Fishing_activity_category_level6) == cs_all@hh$foCatEu6[nr]])
  cs_all@hh$foCatEu6[nr] =   as.character(metier_shortcode$short_code[as.character(metier_shortcode$Fishing_activity_category_level6) == cs_all@hh$foCatEu6[nr]])
   } else {
   print(paste("NOT FOUND: ", cs_all@hh$foCatEu6[nr]))    
   } 
   } 


elements <- c("space",	"time",	"technical", "commercial_category")
  
  col_ele_var1 <<-  as.character(table_strat_res[1,colnames(table_strat_res) == elements[1] ]) 
 col_ele_var2 <<-  as.character(table_strat_res[1,colnames(table_strat_res) == elements[2] ])  
 col_ele_var3 <<-  as.character(table_strat_res[1,colnames(table_strat_res) == elements[3] ]) 
 # col_ele_var4 <<-  as.character(table_strat_res[1,colnames(table_strat_res) == elements[4] ]) 
 

```

Input

*CS COST object*

```{r input1}

head(cs_all@tr)
head(cs_all@hh)
head(cs_all@sl)
head(cs_all@hl)
head(cs_all@ca)

```

*Short codes for metiers*

```{r input2}

head(metier_shortcode)

```

*Stratification table*

```{r input3}

table_strat_res

```


INVESTIGATE DATA FOR THE SELECTED SPECIES

*Calculation of the CV in the past sampling*

```{r data2}
 
 ## ------------------------------------------ ##
 ## INVESTIGATE DATA FOR THE TARGET SPECIES
 ## ------------------------------------------ ##

res_dir <- paste(results_path, "/04_dataInvestigation", sep="")

dir.create(res_dir)

for (this_spe in 1:length( target_SPECIES)) {
  sppName <-  target_SPECIES[[this_spe]]
  sppShortCode <- names(target_SPECIES)[this_spe]
    
cl_spe <<-  subset(cl_all, taxon == sppName, link = TRUE) 
cs_spe <<- try(subsetSpp(cs_all, spp == sppName, link = TRUE), silent=TRUE) 

stratification <<- strIni()

for (ye in 1:length(years)) {
 
   print(paste("Investigate data for", years[ye], sppName ,"in progress...") , quote=F)  
  
  cl0 = cl_spe
  cs0 = cs_spe
  
   cs_ye = try(subset(cs0, year == years[ye], table = "tr", link=TRUE)  , silent=TRUE)
  cl_ye = subset(cl0, year == years[ye], link=TRUE) 
 
if(!is.na(cs_ye@hl$spp[1]) ) {
  # % samples
  hsl = merge(cs_ye@hh, cs_ye@hl)
  

 hsl$SampQuarter <-  quarter(hsl$date) 
  hsl$SampSemester <- ifelse(quarter(hsl$date) == 1 | quarter(hsl$date) == 2, 1, 2)
 
 cl_temp <- cl_ye 

 cl_temp@cl$semester <- 1
 cl_temp@cl$semester[cl_temp@cl$quarter %in% c(3, 4)] <- 2
 
 
switch(col_ele_var1,
    GSA={ 
    AggregationList <- list(hsl$area)
    AggregationList_lan <- list(cl_temp@cl$area)
    },
    Country={ 
    AggregationList <- list(hsl$vslFlgCtry)
    AggregationList_lan <- list(cl_temp@cl$vslFlgCtry)
    },
    ALL = { 
    AggregationList <-  list(rep("ALL", length(hsl$area) ) )
    AggregationList_lan <-  list(rep("ALL", length(cl_temp@cl$area) ) ) 
    },
    stop("Enter something that switches me!")
)

switch(col_ele_var2,
    Q = { 
    AggregationList <- c(AggregationList, list(hsl$SampQuarter) )
    AggregationList_lan <- c(AggregationList_lan, list(cl_temp@cl$quarter) )
    },
    S = { 
    AggregationList <- c(AggregationList, list(hsl$SampSemester) )
    AggregationList_lan <- c(AggregationList_lan, list(cl_temp@cl$semester) )
    },
    Y = { 
    AggregationList <- c(AggregationList, list(rep(1, length(hsl$SampSemester) ) ))
    AggregationList_lan <- c(AggregationList_lan, list(rep(1, length(cl_temp@cl$semester) ) ) )
    },
    stop("Enter something that switches me!")
)

switch(col_ele_var3,    
    lev6 = {  
    AggregationList <- c(AggregationList, list(hsl$foCatEu6) )
    AggregationList_lan <- c(AggregationList_lan, list(cl_temp@cl$foCatEu6) )
    },
    lev4 = {  
    AggregationList <- c(AggregationList, list(hsl$foCatEu5) )
     AggregationList_lan <- c(AggregationList_lan, list(cl_temp@cl$foCatEu5) )
    },
    NONE = { 
    
#     if (col_ele_var4 == "N") {
#     AggregationList <- c(AggregationList, list(rep(1, length(hsl$SampSemester) ) ))
#     AggregationList_lan <- c(AggregationList_lan, list(rep(1, length(cl_temp@cl$semester) ) ) )
#     }
    
     print("Sampling with ALL metiers")
    },
    stop("Enter something that switches me!")
)



# switch(col_ele_var4,    
#     Y = {  
#     AggregationList <- c(AggregationList, list(hsl$commCat) )
#      AggregationList_lan <- c(AggregationList_lan, list(cl_temp@cl$commCat) )
#     },
#     N = { 
#     print(paste("Sampling by", col_ele_var3) )
#     },
#     stop("Enter something that switches me!")
# )


 
  tabsamp = aggregate(hsl$lenNum, by = AggregationList, sum)
  # tabsamp
  # remove rows without stratum information
  ind = which(tabsamp$Group.1 == "" | tabsamp$Group.2 == "" | tabsamp$Group.3 == "")
  if(length(ind) != 0){
    tabsamp = tabsamp[-ind,]
  }
  # rates
  tabsamp$sampling = tabsamp$x / sum(tabsamp$x) * 100
  # % landings
  
  
  
  tabland = aggregate(cl_ye@cl$landWt, by = AggregationList_lan, sum)
  # remove rows without stratum information
  ind = which(tabland$Group.1 == "" | tabland$Group.2 == ""  | tabland$Group.3 == "")
  if(length(ind) != 0) {
    tabland = tabland[-ind,]
  }
  # rates
  tabland$landings = tabland$x / sum(tabland$x) * 100

  tab = merge(tabsamp, tabland, by= c("Group.1", "Group.2", "Group.3"), all = TRUE)
  tab$year = years[ye]
  
  ## NA
  tab$sampling[is.na(tab$sampling)] = 0
  tab$landings[is.na(tab$landings)] = 0
  
  
  colnames(tab)[1:3] = c("Var1", "Var2", "Var3")
  
  poinByStr <- tab[1:3]
 poinByStr$sampSize <- -1
 poinByStr$nMeasTotal <- -1 
 poinByStr$MeanNmeasByStr <- -1
 poinByStr$nOtherSp <- -1 
 poinByStr$cv <- -1
  
   for (nr in 1:nrow(tab)) {
      loc_case = tab[nr,1:3]
      
  # subset by area 
  
switch(col_ele_var1,
    GSA={ CSstr1 = try(subsetCOST(cs_ye, area %in% loc_case$Var1, table = "hh"), silent=TRUE)
    },
    Country={ CSstr1 = try(subsetCOST(cs_ye,  vslFlgCtry %in% loc_case$Var1, table = "hh") , silent = TRUE)
    },
    ALL = { CSstr1 = cs_ye
    },
    stop("Enter something that switches me!")
)

    
switch(col_ele_var2,
    Q = { CSstr2 = try(subsetCOST(CSstr1,  quarter(CSstr1@hh$date)  %in% loc_case$Var2, table = "hh"), silent = TRUE)
    },
    S = { CSstr2 = try(subsetCOST(CSstr1,  ifelse(quarter(CSstr1@hh$date) == 1 | quarter(CSstr1@hh$date) == 2, 1, 2) %in% loc_case$Var2, table = "hh"), silent = TRUE)
    },
    Y = { CSstr2 = CSstr1
    },
    stop("Enter something that switches me!")
)

switch(col_ele_var3,    
    lev6 = { CSstr3 = try(subsetCOST(CSstr2,  foCatEu6 %in% loc_case$Var3, table = "hh") , silent = TRUE)
    },
    lev4 = { CSstr3 = try(subsetCOST(CSstr2,  foCatEu5 %in% loc_case$Var3, table = "hh") , silent = TRUE)
    },
    NONE = {  
    
     if (col_ele_var4 == "N") {
        CSstr3 = CSstr2
        }
    },
    stop("Enter something that switches me!")
)

# switch(col_ele_var4,    
#     Y = { CSstr4 = try(subsetCOST(CSstr2,  commCat %in% loc_case$Var4, table = "hh"), silent = TRUE)
#     },
#     N = { CSstr4 = CSstr3
#     },
#     stop("Enter something that switches me!")
# )

    head(CSstr3) 
    
    # sample size
    if(all(!is.na(CSstr3@hl$trpCode))){
      poinByStr$sampSize[nr] = length(unique(CSstr3@hl$trpCode))
    } else {
      poinByStr$sampSize[nr] = 0
      }

    
    # total number of measures
    poinByStr$nMeasTotal[nr] = sum(CSstr3@hl$lenNum, na.rm = TRUE)
    
    # mean number of measures by sample (trip code)
    if (poinByStr$sampSize[nr] != 0) {
      poinByStr$MeanNmeasByStr[nr] = round(poinByStr$nMeasTotal[nr] / poinByStr$sampSize[nr], 0)
    } else {
      poinByStr$MeanNmeasByStr[nr] = 0
      }
    
  
    # cv
    CSvalstr = csDataVal(CSstr3)
    CLvaly = clDataVal(cl_ye)
    CSconsstr = csDataCons(CSvalstr, stratification)
    CLconsy = clDataCons(CLvaly, stratification)
    dbeOutputy = dbeObject(desc = sppName, species = sppName, catchCat = "Lan",
                           strataDesc = stratification, methodDesc = "analytical")
    # compute cv
    raisingy =  try(RaiseLgth(dbeOutputy, CSconsstr, CLconsy,
                              spp = sppName, taxon = sppName, strDesc = stratification), silent = TRUE)
    
    if(class(raisingy) != "try-error"){
      poinByStr$cv[nr] = raisingy@lenNum$DCRcvIndicator
    } else {
      poinByStr$cv[nr] = NA
      }

   } 
    
  if (!exists("tab2") ) {
   tab2 = merge(tab, poinByStr) 
   colnames(tab2) <- c( "Var1", "Var2", "Var3", "x.x", "sampling", "x.y", "landings", "year", "sampSize", "nMeasTotal", "MeanNmeasByStr", "nOtherSp", "cv")
    tab2$Species <- sppName
 } else {
   to_add <-  merge(tab, poinByStr)
 colnames(to_add) <- c( "Var1", "Var2", "Var3", "x.x", "sampling", "x.y", "landings", "year", "sampSize", "nMeasTotal", "MeanNmeasByStr", "nOtherSp", "cv")
 to_add$Species <- sppName
   tab2 = rbind(tab2, to_add )
 }

  
 # if (!exists("pastSituations")) {
   pastSituations = tab2 
   
#print(pastSituations)
 # } else {
 #   pastSituations = rbind(pastSituations, tab2)
 # }
  
}
}

}


```

Past situation

```{r data3}


pastSituations <- pastSituations[, colnames(pastSituations) != "nOtherSp"]
 write.table(pastSituations, file=paste( res_dir, "/Past situation.csv", sep=""), sep=";", row.names=F, na="")
 
pastSituations
 
```

Sampling vs Landings plot from past situation

```{r data4} 
 


 ## ------------------------------------------ ##
 # SAVE GRAPHS from past situation
 ## ------------------------------------------ ## 

for (this_spe in 1:length(target_SPECIES)) {
  
  sppName <- target_SPECIES[[this_spe]]
  sppShortCode <- names(target_SPECIES)[this_spe]

  pastSituations_spe <- pastSituations[pastSituations$Species == sppName , ]
  
   vect_met <- unique(pastSituations_spe$Var3)
   
    for (met in 1:length(vect_met) ) {
   pastSituations_spe_met <-  pastSituations_spe[pastSituations_spe$Var3 == vect_met[met], ]
     
   windows(height=10, width=8)
   
file_name <- paste( res_dir, "/", sppShortCode," - ", vect_met[met], 
                   " - Sampling vs Landing.jpg", sep="") 
plot_title <- paste(sppName , " - ",vect_met[met], " - Sampling vs Landings", sep="") 
   
   ggplot(pastSituations_spe_met) + 
     geom_point(pastSituations_spe_met, mapping=aes(y=sampling, x=landings, col = as.factor(year)), size = 2)+
     geom_abline(intercept = 0, slope = 1, size = 0.2)+
     theme(axis.text.x = element_text(size=7), axis.text.y = element_text(size=7))+
     facet_grid(Var1 ~ Var2, scales='fixed')+
     theme(legend.title=element_blank(), legend.position="top")+
     labs(y = "% Sampling", x = "% Landings")+
     ylim(0, ceiling(max(pastSituations_spe_met$sampling, pastSituations_spe_met$landings)))+
     xlim(0, ceiling(max(pastSituations_spe_met$sampling, pastSituations_spe_met$landings)))  + ggtitle( plot_title)
   ggsave(file_name, last_plot(), width=8, height=10)  
  }
   
 graphics.off()
 
}
```

Landings by year-quarter-foCatEu6

```{r data5} 
 ## ------------------------------------------ ##
 ## SAVE GRAPHS of Landings
 ## ------------------------------------------ ## 

  file_name <- paste(res_dir, "/Landings by year-quarter-foCatEu6.csv", sep="") 

 tab <- aggregate(cl_all@cl$landWt,by=list(cl_all@cl$area, cl_all@cl$taxon, cl_all@cl$quarter, cl_all@cl$year, cl_all@cl$foCatEu6),sum,na.rm=TRUE)
 colnames(tab) <- c("Area", "Species", "Quarter",  "Year" ,  "foCatEu6", "Value")
  write.table(tab, file_name , sep=";", row.names=F, na="")

  print("Landings by year, quarter and foCatEu6 successfully saved!", quote=F)
 print("***************************************************************************") 
 
 
 for (gs in 1:length(gsas)) {
   tab_this_gsa <- tab[tab$Area == gsas[gs],]
   tab_this_gsa$Value <- tab_this_gsa$Value/1000
    # Quarter Year foCatEu6 Prod (kg)
 file_name <- paste( res_dir, "/Landings by year-quarter-foCatEu6 " ,gsas[gs]," .jpg", sep="") 
plot_title <- paste("Landings (tons) by foCatEu6", gsas[gs])
 
 ggplot(tab_this_gsa, aes(x = Year, y = Value, fill=as.factor(Quarter))) + geom_bar(stat="identity", position="dodge") + ggtitle(plot_title) +  facet_grid(Species ~ foCatEu6, scales='fixed')
 
 ggsave(file=file_name, last_plot(), width=10, height=6)
 try(dev.off(), silent=TRUE)
 }

```
 
Landings by year-quarter-foCatEu6

```{r data6}  
 
## ------------------------------------------ ##
 ## SAVE NO of samples from past situation
 ## ------------------------------------------ ## 

 T_final <- data.frame() 

 for (this_spe in 1:length(target_SPECIES)) {
   
  sppName <- target_SPECIES[[this_spe]]
    sppShortCode <- names(target_SPECIES)[this_spe]

  cl_spe <<-  subset(cl_all, taxon == sppName, link = TRUE)
cs_spe <<- try(subsetSpp(cs_all, spp == sppName, link = TRUE) , silent=TRUE)



 hsl <- merge(cs_spe@hh, cs_spe@sl)
 y <- unique(hsl$year)
gsas <- unique(hsl$area)
 
 for (gs in 1:length(gsas)) {
 hslgs <- hsl[which(hsl$area==gsas[gs]),]
   
 for (i in 1:length(y)) {
   hsly <- hslgs[which(hslgs$year==y[i]),]
   
   t_1 <- data.frame(tapply(hsly$commCat, list(hsly$foCatEu6, quarter(hsly$date)), length))
   
   if (nrow(t_1) != 0) {
      colnames(t_1) <- unique(quarter(hsly$date))[order(unique(quarter(hsly$date)))]
   rownames(t_1) <- paste(as.character(y[i]),rownames(t_1))
   
   to_add <- which(!c(1:4) %in%  unique(quarter(hsly$date)))
   
   if (length(to_add) != 0) {     
     ad <- data.frame(matrix(0, ncol=length(to_add), nrow=nrow(t_1)))
     colnames(ad) <- to_add
     t_1 <- cbind(t_1,ad)
     t_1 <- t_1[, order(colnames(t_1))]
   }   

      t_1 <- data.frame(cbind(as.character(gsas[gs]), cbind(as.character(y[i]), cbind( substring(rownames(t_1), 6, nchar(rownames(t_1))), cbind(sppName, t_1)))))
   colnames(t_1)[1:4] <- c("Area", "Year", "Metier", "Species")
   
   T_final <- rbind(T_final, t_1)
   
     
   }
  
 } # close area
  
 } #close year 

 } # close species

 tab1 <- data.frame(Area=T_final$Area,Year=T_final$Year, Metier=T_final$Metier, Species=T_final$Species, SQ1=T_final[,5], SQ2=T_final[,6], SQ3=T_final[,7],  SQ4=T_final[,8])

## ------------------------------------------ ##
 ## SAVE NO of individuals from past situation
 ## ------------------------------------------ ## 
 
  T_final2 <- data.frame() 

 for (this_spe in 1:length(target_SPECIES)) {
  sppName <- target_SPECIES[[this_spe]]
    sppShortCode <- names(target_SPECIES)[this_spe]

  cl_spe <<-  subset(cl_all, taxon == sppName, link = TRUE)
cs_spe <<- try( subsetSpp(cs_all, spp == sppName, link = TRUE), silent=TRUE) 

 hsl2 <- merge(cs_spe@hh, cs_spe@hl)
 
 y <- unique(hsl2$year)
gsas <- unique(hsl2$area)
 
 for (gs in 1:length(gsas)) {
 hslgs <- hsl2[which(hsl2$area==gsas[gs]),]
   
 for (i in 1:length(y)) {
   hsl2y <- hslgs[which(hslgs$year==y[i]),]
   
   t_2 <- data.frame(tapply(hsl2y$lenNum, list(hsl2y$foCatEu6, quarter(hsl2y$date)), sum))
   
      if (nrow(t_2) != 0) {
   colnames(t_2) <- unique(quarter(hsl2y$date))[order(unique(quarter(hsl2y$date)))]
   rownames(t_2) <- paste(as.character(y[i]),rownames(t_2))
   to_add <- which(!c(1:4) %in%  unique(quarter(hsl2y$date)))
   if (length(to_add) != 0) {     
     ad <- data.frame(matrix(0, ncol=length(to_add), nrow=nrow(t_2)))
     colnames(ad) <- to_add
     t_2 <- cbind(t_2,ad)
     t_2 <- t_2[, order(colnames(t_2))]
   }
   
 t_2 <- data.frame(cbind(as.character(gsas[gs]), cbind(as.character(y[i]), cbind( substring(rownames(t_2), 6, nchar(rownames(t_2))), cbind(sppName, t_2)))))
   colnames(t_2)[1:4] <- c("Area", "Year", "Metier", "Species")
   
   T_final2 <- rbind(T_final2,t_2)
      }
 } # close area
 
   } # close year
 
 } # close species

 tab2 <- data.frame(Area=T_final2$Area, Year=T_final2$Year, Metier=T_final2$Metier, Species=T_final2$Species, NQ1=T_final2[,5], NQ2=T_final2[,6], NQ3=T_final2[,7],  NQ4=T_final2[,8])
 
 ## ------------------------------------------ ##
 ## SAVE NO of trips from past situation
 ## ------------------------------------------ ## 
 
  tab_TRIP <- tab2

 no_trips_mat <- cs_all@hh
 
 no_trips_mat$ quarter <-  quarter(no_trips_mat$date)
 
 no_trips_mat <- with(no_trips_mat, aggregate(year, by =list(area, year, foCatEu6, quarter, trpCode ), FUN="length") )
 colnames(no_trips_mat) <- c("area", "year", "foCatEu6", "quarter", "trpCode", "x")
 
 no_trips_mat_2 <- with(no_trips_mat, aggregate(year, by =list(area, year, foCatEu6, quarter ), FUN="length") )
 colnames(no_trips_mat_2) <- c("area", "year", "foCatEu6", "quarter", "no_trips")
 
 for (nr in 1:nrow(no_trips_mat_2)) {
   if (no_trips_mat_2$ quarter[nr] == 1) {
     tab_TRIP[tab_TRIP$Area == no_trips_mat_2$area[nr] & tab_TRIP$Year == no_trips_mat_2$year[nr] & tab_TRIP$Metier == no_trips_mat_2$foCatEu6[nr], 5] <- no_trips_mat_2$no_trips[nr]
   } else if (no_trips_mat_2$ quarter[nr] == 2) {
     tab_TRIP[tab_TRIP$Area == no_trips_mat_2$area[nr] & tab_TRIP$Year == no_trips_mat_2$year[nr] & tab_TRIP$Metier == no_trips_mat_2$foCatEu6[nr], 6] <- no_trips_mat_2$no_trips[nr]
   } else if (no_trips_mat_2$ quarter[nr] == 3) {
     tab_TRIP[tab_TRIP$Area == no_trips_mat_2$area[nr] & tab_TRIP$Year == no_trips_mat_2$year[nr] & tab_TRIP$Metier == no_trips_mat_2$foCatEu6[nr], 7] <- no_trips_mat_2$no_trips[nr]
   } else if (no_trips_mat_2$ quarter[nr] == 4) {
     tab_TRIP[tab_TRIP$Area == no_trips_mat_2$area[nr] & tab_TRIP$Year == no_trips_mat_2$year[nr] & tab_TRIP$Metier == no_trips_mat_2$foCatEu6[nr], 8] <- no_trips_mat_2$no_trips[nr]
   }
 }
 
 colnames(tab_TRIP) <- c("Area", "Year", "Metier", "Species", "TQ1", "TQ2", "TQ3", "TQ4")
 tab0 <- merge(tab_TRIP, merge(tab1,tab2))
 index_ <- which(is.na(tab0[,]) , arr.ind = TRUE)
 
 if (nrow(index_) > 0 ){
   for (ind in 1:nrow(index_)) {
     tab0[index_[ind,1], index_[ind,2]] <- 0
   }
 }
 
 tab0[,5:16] <- round( tab0[,5:16],0)
 
 tab0$annualT <- rowSums(tab0[, 5:8])
 tab0$annualS <- rowSums(tab0[, 9:12])
 tab0$annualN <- rowSums(tab0[, 13:16])
 
 file_name <- paste( res_dir, "/No. of trips-samples-measures.csv", sep="") 
 write.table(tab0,  file_name, row.names=F, sep=";", na="")

 print("Trips, samples and measures for the past successfully saved!", quote=F)
 print("***************************************************************************") 

```  
 
 Weight by trip from past situation
 
```{r data7}
 ## ------------------------------------------ ##
 ## SAVE graphs of weight by trip from past situation
 ## ------------------------------------------ ## 
 
 for (this_spe in 1:length(target_SPECIES) ) {
  sppName <- target_SPECIES[[this_spe]]
    sppShortCode <- names(target_SPECIES)[this_spe]

  cs_spe <<- try(subsetSpp(cs_all, spp == sppName, link = TRUE) , silent=TRUE)
  
#  cs_spe@hl$landCat <- "HUC"
#  cs_spe@hl$commCatScl <- "EU"
#  
#  #CSsp@sl$commCat  <- ""

#  CSsp@sl$sex <- ""
#  
#  #CSsp@hl$commCat  <- ""

 
for (gs in gsas) { 

 for (ye in years) {
    print(paste("Saving weight by trip for", sppName, gs, ye, "..."))
   # We start to look at the sample weight by trip, metier and quarter. 
   ##=============================##
   
plot_path <- paste( res_dir, "/",sppShortCode," - Sample weight by trip-metier-quarter ", ye," - ", gs, ".jpg", sep="") 
     
    myCSsub <- try(subset(cs_spe, year == ye, table="sl", link=TRUE), silent=TRUE) # 
    myCSsub <- try(subset(myCSsub,  area == gs , table="hh", link=TRUE) , silent=TRUE) # 

    if(!is.na(myCSsub@sl$spp[1]) ) {
    
    unique(myCSsub@hh$year)
   unique(myCSsub@hh$area)
   unique(myCSsub@sl$spp)
   
  # myCSsub_temp =myCSsub
   
   #strDates <- myCSsub@hh$date
   #dates <- as.Date(strDates, "%d/%m/%Y")
   #myCSsub@hh$date <- dates
   
   strD <- strIni( timeStrata="quarter", techStrata="foCatEu6", spaceStrata = "area")     
   LDVpd <- landisVol(myCSsub,strDef = strD, species=sppName  ) # 
   
  # jpeg(file=plot_path, width=36, height=21, bg="white", units="cm",res=200)
   
   #par(mar=c(8, 5, 5, 5)) 
   #boxplot(LDVpd, cex.lab=0.5, rot=20)
   #try(dev.off() , silent=TRUE)
    
   #myCSsub <- myCSsub_temp
     
plot_path <- paste( res_dir, "/",sppShortCode," - Consistency sampling vs landings ", ye," - ", gs, ".jpg", sep="") 
   
   myCLsub <- subset(cl_all,year==ye & taxon==sppName & area == gs, table="cl")
   #myCLsub@cl$taxon <- sppName
   
   ## VALIDATION
   ## CS
   csVal <- csDataVal(myCSsub)
   ## CL
   clVal <- clDataVal(myCLsub)
   ## STRATIFICATION
   strD <- strIni(timeStrata="quarter",techStrata="foCatEu6")
   
   ## Check consistency of sampling/prod
   CSrel <- relativeValue(csVal, strD, "nbSamp")
   CLrel <- relativeValue(clVal, strD)
   
   jpeg(file=plot_path, width=21, height=27, bg="white", units="cm",res=200)
   plot(CLrel,CSrel)
   try(dev.off(), silent=TRUE)
 
    }
   
   } # close year
  

} # close gsa

}
 
```
 
 Investigating variability between samples
 
```{r data8}
 ## ------------------------------------------ ##
 ## Investigating variability between samples
 ## ------------------------------------------ ## 

  for (this_spe in 1:length(target_SPECIES) ) {
  sppName <- target_SPECIES[[this_spe]]
    sppShortCode <- names(target_SPECIES)[this_spe]

  cs_spe <<- try(subsetSpp(cs_all, spp == sppName, link = TRUE) , silent=TRUE)
  
for (gs in gsas) { 
     
myCSsub <- try( subset(cs_spe,  area == gs , table="hh") , silent=TRUE)

   if (!is.na(myCSsub@sl$sampType[1]))  {

plot_path <- paste( res_dir, "/",sppShortCode," - Variability between samples by year - ", gs, ".jpg", sep="") 
delta_path <- paste( res_dir, "/",sppShortCode," - Variability between samples by year - ", gs, ".csv", sep="") 
   
test_hl <- myCSsub@hl 

test_2 <- group_by(test_hl, sampType, year, trpCode, staNum, spp, commCat, sex)
test_2 <- summarize(test_2, no_lengths= length(sampType))

 strD <- strIni(timeStrata="year")

 deltas <- deltCalc(myCSsub, species=sppName, strDef=strD, method="subsample", strategy="metier")
 
 jpeg(file=plot_path, width=36, height=21, bg="white", units="cm",res=200)
 par(mar=c(8, 5, 5, 5))                                                                                                                                     
 plot(deltas)  
 try(dev.off(), silent=TRUE )
 
 delta_res <- merge(deltas@ outPut$ SampDeltaMat, deltas@ outPut$ DFsamp, by.x=c("samp"), by.y=c("SampNum") )
 write.table(delta_res, file=delta_path, row.names=F, sep=";", na="")

 plot_path <- paste( res_dir, "/",sppShortCode," - Variability between samples by metier - ", gs, ".jpg", sep="") 
delta_path <- paste( res_dir, "/",sppShortCode," - Variability between samples by metier - ", gs, ".csv", sep="") 

 strD <- strIni(techStrata="foCatEu6")
 deltas <- deltCalc(myCSsub,species=sppName,strDef=strD, method="subsample", strategy="metier")
 
 jpeg(file=plot_path, width=36, height=21, bg="white", units="cm",res=200)
 par(mar=c(8, 5, 5, 5))       
 plot(deltas) 
 try(dev.off(), silent=TRUE)
 
 # method="subsample"
 # Method for calculating delta values. If "delta", samples are raised to haul/trip level. 
 # If "subsample", samples are considered at sampling level (no raising process).
 
 delta_res <- merge(deltas@ outPut$ SampDeltaMat, deltas@ outPut$ DFsamp, by.x=c("samp"), by.y=c("SampNum") )
 write.table(delta_res, file=delta_path, row.names=F, sep=";", na="")
 
plot_path <- paste( res_dir, "/",sppShortCode," - Variability between samples by quarter - ", gs, ".jpg", sep="") 
delta_path <- paste( res_dir, "/",sppShortCode," - Variability between samples by quarter - ", gs, ".csv", sep="") 
 
 strD <- strIni(timeStrata="quarter")
 
  myCSsub_temp =myCSsub
   
strDates <- myCSsub@hh$date
 dates <- as.Date(strDates, "%d/%m/%Y")
 myCSsub@hh$date <- dates
 
 deltas <- deltCalc(myCSsub,species=sppName,strDef=strD, method="subsample", strategy="metier")
 jpeg(file=plot_path, width=36, height=21, bg="white", units="cm",res=200)
 par(mar=c(8, 5, 5, 5))   
 plot(deltas)  
 try(dev.off() , silent=TRUE)
 
myCSsub = myCSsub_temp 
 
 delta_res <- merge(deltas@ outPut$ SampDeltaMat, deltas@ outPut$ DFsamp, by.x=c("samp"), by.y=c("SampNum") )
 write.table(delta_res, file=delta_path, row.names=F, sep=";", na="")

   }


  for (anno in years) {

    plot_path <- paste( res_dir, "/",sppShortCode," - Consistency sampling vs landings ", anno," - ", gs, ".jpg", sep="") 
    
   myCSsub <- subset(cs_spe , year==anno & area==gs, table="hh")
   myCLsub <- subset(cl_spe, year==anno & area==gs, table="cl")
   
   unique(myCSsub@hh$year)
   unique(myCLsub@cl$year)
   
   unique(myCSsub@hh$area)
   unique(myCLsub@cl$area)
   
   unique(myCSsub@sl$spp)
   unique(myCLsub@cl$taxon)
   
      if (!is.na(myCSsub@sl$spp[1]))  {
   ## VALIDATION
   ## CS
   csVal <- csDataVal(myCSsub)
   ## CL
   clVal <- clDataVal(myCLsub)
   ## STRATIFICATION
   strD <- strIni(timeStrata="quarter",techStrata="foCatEu6")
   ## Check consistency of sampling/prod
   CSrel <- relativeValue(csVal, strD, "nbSamp")
   CLrel <- relativeValue(clVal, strD)

   jpeg(file=plot_path, width=21, height=27, bg="white", units="cm",res=200)
   plot(CLrel,CSrel)
   try(dev.off() , silent=TRUE)
      }
   
  }
 
}
 }
 
 
```
 
 Save lenght distribution plot by trip
 
```{r data9}
  # ---------------------------------------------------------------- save lenght distribution plot by trip
 
  for (this_spe in 1:length(target_SPECIES) ) {
  sppName <- target_SPECIES[[this_spe]]
    sppShortCode <- names(target_SPECIES)[this_spe]

  cs_spe <<- subsetSpp(cs_all, spp == sppName, link = TRUE) 
  
 for (gs in gsas) { 
     
    for (anno in years) {
   
 plot_path <- paste( res_dir, "/",sppShortCode," - Length distribution by trip ", anno," - ", gs, ".jpg", sep="") 
    
   myCSsub <- subset(cs_spe , year==anno & area==gs, table="hh")
   
     if (!is.na(myCSsub@sl$sampType[1]))  {
    unique(myCSsub@hh$year)
   unique(myCSsub@hh$area)
   unique(myCSsub@sl$spp)

   # plot length distribution 
   jpeg(file=plot_path, width=46, height=21, bg="white", units="cm",res=200)
   lenDisPlot(myCSsub,sppName,fraction="LAN",trpCode="all",level="trip")  
   try(dev.off() , silent=TRUE)
     }
   
 }
 
 
 }


} # close species
 

```
